Designing a modular architecture for building web-based psychological experiments

### 2021-08-24 paper.tex
Fix typos

\chapter{Introduction}
Web-based experiments are gaining more popularity in behavioural research because of their flexibility and efficiency. I conducted an analsysis of prior art and found that many existing tools for realizing web-based experiments lack sufficient flexibility and expressiveness. This lead to the creation of magpie, both a server for managing experiments and a client-side framework for implementing them.

@@

In this work, I initially examine prior work and competing software products with a set of requirements in mind. In the following section I explore the design goals and choices in relation to these requirements that lead to the implementation of magpie's client-side framework, its concepts, components, features and technical realization. Afterwards, I discuss limitations of the current design and directions for furture work. Finally, I reach conclusions about the design and its applications.

@@

\textbf{psiTurk} \citep{Gureckis2016} is mainly a server-side framework for integrating web-based experiments with Amazon Mechanical Turk and cannot be used with other recruiting platforms. It only aids minimally in developing the actual experiments and thus has to be combined with a different front-end framwork.

@@

\begin{table}[]
\begin{tabular}{lllllllllll}
                      & \rot{OpenSesame} & \rot{PsyToolkit} & \rot{psiTurk}    & \rot{jsPsych}    & \rot{LIONESS-lab} & \rot{Labvanced}  & \rot{zTree}      & \rot{Dallinger}  & \rot{lab.js}     & \rot{magpie}     \\
Web                        & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark &            & \checkmark & \checkmark & \checkmark \\
Free software              & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  &            & \checkmark & \checkmark & \checkmark & \checkmark \\
Reruiting flexibility      & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark &            &            &            & \checkmark \\
Sharable customizations    &            &            &            & \checkmark &             & ?          & ?          &            & \checkmark & \checkmark \\
Framework                  & \checkmark & \checkmark &            &            & \checkmark  & \checkmark & \checkmark &            & \checkmark & \checkmark \\
Catalogue                  & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark & \checkmark &            &            & \checkmark \\
Interactivity              &            &            &            &            & \checkmark  & \checkmark & \checkmark & \checkmark &            & \checkmark \\
Dynamicity                 & \checkmark &            &            &            & \checkmark  & ?          & ?          & ?          & ?          & \checkmark \\
Multimedia                 & \checkmark & \checkmark          &            & \checkmark & \checkmark  & \checkmark & ?          &            &            & \checkmark \\
Implementation flexibility & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark &            &            & \checkmark & \checkmark
\end{tabular}
\caption{Overview of features}
\label{tab:competition}
\end{table}

@@

To this end, taking advantage of Vue.js allows for fast progress and interoperability with a large catalogue of open source modules already built on that same framework, while avoiding mistakes that others have already made and corrected. Choosing a reactive framework is a major benefit for implementation effiency, because the view layer updates automatically based on the provided data, reducing the amount of necessary code.

@@

These finer grained components offer essential features that can be recombined in a variety of ways to form trials of any complexity.
To ease the building of experiments, there are many pre-built screen components available. These allow the quick assembly of simple experiments out of coarse building blocks, which utilize the above fine graned building blocks under the hood. These trial screens follow a life cycle of five slides: pause, fixation, stimulus presentation, response collection and feedback. All phases except response collection are optional, and all phases can be configured with a timeout or a manual trigger to go to the next phase. Additionally, implementers can add textual context like a Question. There are also slots available for each of the four life cycle phases, that can be filled with custom content.

@@

The screen will first display a fixation cross for a random duration between 1200ms and 2700ms, specified by the \texttt{fixation-time} attribute. Afterwards, the simulus is displayed allowing a response with the Q and P keys. The response is stored in \texttt{\$magpie.measurements.response}. The stimuli are displayed using the built-in canvas components, displaying the square or circle left or right of the screen. The Record component adds the trial data to the result measurements. The \texttt{response-time} attribute sets a timeout of three seconds to respond, after which the participant is shown the feedback screen which tells them to respond more quickly if they haven't responded yet.

>>>>

\chapter{Introduction}
Web-based experiments are gaining more popularity in behavioural research because of their flexibility and efficiency. I conducted an analysis of prior art and found that many existing tools for realizing web-based experiments lack sufficient flexibility and expressiveness. This lead to the creation of magpie, both a server for managing experiments and a client-side framework for implementing them.

@@

In this work, I initially examine prior work and competing software products with a set of requirements in mind. In the following section I explore the design goals and choices in relation to these requirements that lead to the implementation of magpie's client-side framework, its concepts, components, features and technical realization. Afterwards, I discuss limitations of the current design and directions for future work. Finally, I reach conclusions about the design and its applications.

@@

\textbf{psiTurk} \citep{Gureckis2016} is mainly a server-side framework for integrating web-based experiments with Amazon Mechanical Turk and cannot be used with other recruiting platforms. It only aids minimally in developing the actual experiments and thus has to be combined with a different front-end framework.

@@

\begin{table}[]
\begin{tabular}{lllllllllll}
                      & \rot{OpenSesame} & \rot{PsyToolkit} & \rot{psiTurk}    & \rot{jsPsych}    & \rot{LIONESS-lab} & \rot{Labvanced}  & \rot{zTree}      & \rot{Dallinger}  & \rot{lab.js}     & \rot{magpie}     \\
Web                        & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark &            & \checkmark & \checkmark & \checkmark \\
Free software              & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  &            & \checkmark & \checkmark & \checkmark & \checkmark \\
Recruiting flexibility      & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark &            &            &            & \checkmark \\
Sharable customizations    &            &            &            & \checkmark &             & ?          & ?          &            & \checkmark & \checkmark \\
Framework                  & \checkmark & \checkmark &            &            & \checkmark  & \checkmark & \checkmark &            & \checkmark & \checkmark \\
Catalogue                  & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark & \checkmark &            &            & \checkmark \\
Interactivity              &            &            &            &            & \checkmark  & \checkmark & \checkmark & \checkmark &            & \checkmark \\
Dynamicity                 & \checkmark &            &            &            & \checkmark  & ?          & ?          & ?          & ?          & \checkmark \\
Multimedia                 & \checkmark & \checkmark          &            & \checkmark & \checkmark  & \checkmark & ?          &            &            & \checkmark \\
Implementation flexibility & \checkmark & \checkmark &            & \checkmark & \checkmark  & \checkmark &            &            & \checkmark & \checkmark
\end{tabular}
\caption{Overview of features}
\label{tab:competition}
\end{table}

@@

To this end, taking advantage of Vue.js allows for fast progress and interoperability with a large catalogue of open source modules already built on that same framework, while avoiding mistakes that others have already made and corrected. Choosing a reactive framework is a major benefit for implementation efficiency, because the view layer updates automatically based on the provided data, reducing the amount of necessary code.

@@

These finer grained components offer essential features that can be recombined in a variety of ways to form trials of any complexity.
To ease the building of experiments, there are many pre-built screen components available. These allow the quick assembly of simple experiments out of coarse building blocks, which utilize the above fine graned building blocks under the hood. These trial screens follow a life cycle of five slides: pause, fixation, stimulus presentation, response collection and feedback. All phases except response collection are optional, and all phases can be configured with a timeout or a manual trigger to go to the next phase. Additionally, implementers can add textual context like a Question. There are also slots available for each of the four life cycle phases that can be filled with custom content.

@@

The screen will first display a fixation cross for a random duration between 1200ms and 2700ms, specified by the \texttt{fixation-time} attribute. Afterwards, the stimulus is displayed allowing a response with the Q and P keys. The response is stored in \texttt{\$magpie.measurements.response}. The stimuli are displayed using the built-in canvas components, displaying the square or circle left or right of the screen. The Record component adds the trial data to the result measurements. The \texttt{response-time} attribute sets a timeout of three seconds to respond, after which the participant is shown the feedback screen which tells them to respond more quickly if they haven't responded yet.
